### 2022-12-09 introduction/introduction.tex
Added cryo-EM as example of structural technique

Because sequence dictates structure and structure dictates function, understanding these relationships in proteins is essential for understanding biological systems as a whole. While biologists are generally interested in function since it is the most directly associated with the biological role\footnote{Function and biological role are frequently, though not always, exact synonyms since biological role is often defined more broadly and loosely than function. For example, the strict function of hemoglobin is to bind oxygen, an activity it shares with a related protein myoglobin. However, the two have different roles. Hemoglobin is found in red blood cells where it primarily acts a carrier. In contrast, myoglobin is found in muscle cells, where it stores oxygen until needed.} of a protein, function and biological role are not always easily identified or measured. Thus, determining a protein's structure is frequently the first step of detailed study of a protein's function. Though in recent years researchers have developed powerful computational tools that can accurately predict structure from sequence alone, historically structures were determined experimentally, and experimental methods remain the gold standard today. While many methods can reveal information about the structure of a protein, the most powerful techniques, X-ray crystallography and NMR spectroscopy, can map the spatial coordinates of every atom in a protein. However, this resolution requires extremely pure samples of the protein of interest. Since proteins are only produced by living systems, preparations begin with a complex mixture consisting of cells or tissue, and the protein of interest is isolated with series of extraction and purification steps. Some proteins are only produced in small amounts or degrade easily, so each step may require substantial optimization to achieve a sufficient yield. When structural techniques were first developed in the late 1950s, they were so time-consuming that a graduate student could dedicate an entire PhD to solving a single protein structure. Many developments have substantially accelerated the process, but it remains a labor-intensive technique which may require several months of effort. However, the result is a powerful map that scientists use to suggest hypotheses and interpret data.

>>>>

Because sequence dictates structure and structure dictates function, understanding these relationships in proteins is essential for understanding biological systems as a whole. While biologists are generally interested in function since it is the most directly associated with the biological role\footnote{Function and biological role are frequently, though not always, exact synonyms since biological role is often defined more broadly and loosely than function. For example, the strict function of hemoglobin is to bind oxygen, an activity it shares with a related protein myoglobin. However, the two have different roles. Hemoglobin is found in red blood cells where it primarily acts a carrier. In contrast, myoglobin is found in muscle cells, where it stores oxygen until needed.} of a protein, function and biological role are not always easily identified or measured. Thus, determining a protein's structure is frequently the first step of detailed study of a protein's function. Though in recent years researchers have developed powerful computational tools that can accurately predict structure from sequence alone, historically structures were determined experimentally, and experimental methods remain the gold standard today. While many methods can reveal information about the structure of a protein, the most powerful techniques, X-ray crystallography, NMR spectroscopy, and cryogenic electron microscopy (cryo-EM), can map the spatial coordinates of every atom in a protein. However, this resolution requires extremely pure samples of the protein of interest. Since proteins are only produced by living systems, preparations begin with a complex mixture consisting of cells or tissue, and the protein of interest is isolated with series of extraction and purification steps. Some proteins are only produced in small amounts or degrade easily, so each step may require substantial optimization to achieve a sufficient yield. When structural techniques were first developed in the late 1950s, they were so time-consuming that a graduate student could dedicate an entire PhD to solving a single protein structure. Many developments have substantially accelerated the process, but it remains a labor-intensive technique which may require several months of effort. However, the result is a powerful map that scientists use to suggest hypotheses and interpret data.
